An Infinite de Bruijn Sequence

### 2016-12-22 infdebruijn.tex
With Gal's section

\begin{proposition}[Prefer Maximum]
For every $\sigma \in \N$ and $w \in \N^{n-1}$, the word $\sigma w$ comes before the word $(\sigma+1) w$ in $C_\infty$.
\end{proposition}
\begin{proof}
Let $k$ be a key word  of the form $0^z(k)\hat{w}$. We will prove the proposition for every rotation of $k$:

\begin{description} 
\item [First Case:] If $w=0^{z_1}\hat{w}0^{z_2}$ such that $z_1+z_2=z(k)$ and $z_1 \neq 0$ then we have that $w$ precedes  $w_1=0^{z_1-1}\hat{w}0^{z_2}1$ by the definition of $nxt$ and that $w_1$ precedes $1 0^{z_1-1}\hat{w}0^{z_2}$ by the order that we added the words in this necklace.
\item [Second Case:] If $w=\sigma u 0^{z(k)}v$ such that $v\sigma u=\hat{w}$ we have that
\begin{itemize}
	\item $w$ precedes $v \sigma u 0^{z(k)}$ because the latter is last on its necklace.
	\item $v \sigma u 0^{z(k)}$ precedes $\sigma u 0^{z(k)+|v|}$ because...
	\item $\sigma u 0^{z(k)+|v|}$ precedes $0(\sigma+1) u 0^{z(k)+|v|-1}$ because...
	\item $0(\sigma+1) u 0^{z(k)+|v|-1}$ precedes the word $u_2 (\sigma+1) u_1$ such that $u_1u_2=u0^{z(k)+|v|+1}$ and $last(u_2 (\sigma+1) u_1)$, by the order of the enumeration of the words in this necklace.
	\item $u_2 (\sigma+1) u_1$ precedes $0(\sigma+1)u_1 0^{|u_2|-1}$ because...
	\item $0(\sigma+1)u_1 0^{|u_2|-1}$ precedes $0 \tau_2 \dots \tau_l (\sigma+1)u_1$ where $u_2=\tau_1\dots\tau_l \sigma u_1=v\sigma u 0^{z(k)}$ because... 
	\item $0 \tau_2 \dots \tau_l (\sigma+1)u_1$ precedes $\tau_2 \dots \tau_l (\sigma+1)u_1\tau_1$ because... 
	\item $\tau_2 \dots \tau_l (\sigma+1)u_1\tau_1$ precedes $(\sigma+1)u_1\tau_1 \tau_2 \dots \tau_l=(\sigma+1) u 0^{z(k)}v$ because... 
\end{itemize}
\end{description}

@@

\begin{theorem}
$w^\infty$ is an infinite De-Bruijn sequence.
\end{theorem}

@@

>>>>

\begin{proposition}[Prefer Maximum]
For every $\sigma \in \N$ and $w \in \N^{n-1}$, the word $\sigma w$ comes before the word $(\sigma+1) w$ in $C_\infty$.
\end{proposition}

@@

\begin{theorem}
	$w^\infty$ is an infinite de Bruijn sequence.
\end{theorem}

@@

\section{Constructing an Infinite de Bruijn Cycle}

A key word is a word of length $n$ that is maximal in right-to-left (Arabic) lexicographic order among its cyclic rotations. Let $kw_0,kw_1,kw_2,\dots$ be an enumeration of all key-words, ordered in right-to-left (Arabic) lexicographic order. Let $C_m=\{R^i(kw_m): i=1,\dots,n\}$ be the cycle of $kw_m$. We order the elements of $C_m$ as follows:  if $kw_m=0^l(\sigma+1)w$, then $w0^l(\sigma+1)$ is the first word in $C_m$, and each word $w'$ is followed by $R(w')$. The last word in $C_m$ is $(\sigma+1)w0^l$.\\

For each $m<n$ we define a de Bruijn sequence $D_m$, over the words $\bigcup_{i=0}^m C_i$ as follows: 
\begin{itemize}
	\item $D_0=0^n$.
	\item If $kw_{m+1}=0^l(\sigma+1)w$, then $D_{m+1}$ is obtained by inserting the sequence $C_{m+1}$ after the word $\sigma w 0^l\in C_m$.
\end{itemize}

%\begin{lemma}
%Every $D_m$ is a de Bruijn sequence over $\bigcup_{i=0}^m C_i$, and   $\forall m$ $\exists k$ so that if $k\leq r$, then $D_m$ is a prefix of $D_r$. 
%\end{lemma}

\begin{definition}
	For $w,w\in \N^n$ and $m\in \N$, write $w<_m w'$ if $w$ appears before $w'$ in $D_m$. Write $<=\bigcup_{i=0}^\infty$.
\end{definition}

For a word $w$, $max(w)$ is the maximal number in $w$.

\begin{lemma}
	If $w<w'$, then $max(w)\leq max(w')$.
\end{lemma}

\begin{corrolary}
	$<$ defines an infinite de Bruijn sequence.
\end{corrolary}

\begin{proof}
	By the previous Lemma, each word is preceded by finitely many words thus $<$ defines an infinite sequence. Since each $D_m$ is a de Bruijn sequence and since $<_m\subseteq<_{m+1}$, the sequence defined by $<$ is a de Bruijn sequence. 
\end{proof}

Let $D$ be the infinite de Bruijn sequence defined by $<$.

\begin{theorem}
	For a  word $\sigma w$ in $D$, let $nxt(\sigma w)$ be the successor of $\sigma w$ in $D$. Then,
	$$nxt(\sigma w)=\left\{ \begin{array}{lr}
	w(\sigma+1), & last((\sigma+1)w)\\
	w0, & last(\sigma w)\\
	w\sigma, & otherwise
	\end{array}\right.$$
\end{theorem}

\begin{definition}
	If $kw_m=0^l(\sigma+1)w$, we write $first(w0^l(\sigma+1))$, $key(0^l(\sigma+1)w)$ and $last((\sigma+1)w0^l)$. In addition, for a cycle $C_m$, $first(C_m)=w\in C_m$ so that $first(w)$,  $key(C_m)=kw_m$ and $last(C_m)=w\in C_m$ so that $last(w)$. 
\end{definition}

\begin{lemma}
	For any $w'\in C_m$, $first(C_m)\leq w'\leq last(C_m)$.
\end{lemma}
\begin{proof}
	This the way we ordered the cycles.
\end{proof}

\begin{definition}
	We say that $C_{i_0}C_{i_1}\dots C_{i_k}$ is a sequence of cycles, if for every $j<k$, $nxt(last(C_{i_j}))=first(C_{i_{j+1}})$.
\end{definition}

\begin{lemma}
	\label{sequence}
	Let $C_{i_0}C_{i_1}\dots C_{i_k}$ be a sequence of cycles. Write, $key(C_{i_0})=0^l(\sigma)w$ where $\sigma\neq 0$. Then, for every $j\leq k$, $key(C_{i_j})=0^l(\sigma+j)w$.
\end{lemma}

\begin{proof}
	Assume by induction that $key(C_{i_j})=0^l(\sigma+j)w$ for $j<k$. Hence, $last(C_{i_j})=(\sigma+j)w0^l$. Thus, $nxt((\sigma+j)w0^l)=w0^l(\sigma+j+1)$ or $nxt((\sigma+j)w0^l)=w0^{l+1}$. Since $\neg(first(w0^{l+1}))$, we have 
	$$nxt((\sigma+j)w0^l)=w0^l(\sigma+j+1)=first(C_{i_{j+1}}).$$
	We conclude that $key(C_{i_{j+1}})=0^l(\sigma+j+1)w$.
\end{proof}

\begin{lemma}[The parentheses property]
	\label{parenthesis}
	For any two cycles $C_k$ and $C_m$, one of the following occur
	\begin{itemize}
		\item $last(C_k)<first(C_m)$ or $last(C_m)<first(C_k)$.

		\item $first(C_k)<first(C_m)\leq last(C_m)<last(C_k)$ or \\
		$first(C_m)<first(C_k)\leq last(C_k)<last(C_m)$.
	\end{itemize}
\end{lemma}
\begin{proof}
	This concluded by the way $D_{m+1}$ is obtained from $D_m$.
\end{proof}

\begin{definition}
	We say that $C_m$ is embedded in $C_k$ if $first(C_k)<first(C_m)\leq last(C_m)<last(C_k)$.\\ 
	In addition, $C_m$ is said to be immediately embedded in $C_k$ if there is no $C_l$ such that $C_m$ is embedded in $C_l$ and $C_l$ is embedded in $C_k$.\\ 
	We define by inductively the statement: ``$C_m$ is $r$-embedded in $C_k$":
	\begin{itemize}
		\item $C_m$ is $1$-embedded in $C_k$ if it is immediately embedded in $C_k$.
		\item $C_m$ is $r$-embedded in $C_k$ if there is a cycle $C_l$ such that $C_m$ is $r-1$-embedded in $C_l$ and $C_l$ is immediately embedded in $C_k$. 
	\end{itemize} 
\end{definition}

\begin{lemma}
	\label{embedding-properties}
	Assume that $C_m$ is immediately embedded $C_k$. Write $key(C_m)=0^i(\sigma+1)0^jw$ where $w$ does not starts with $0$. Then,
	\begin{itemize}
		\item $key(C_k) =0^{i+1+j}w$.
		\item If $u\in C_k$ and $last(C_m)<u$, then $u=0^{j_2}w0^{i+1+j_1}$ where $j_1+j_2=j$. 
	\end{itemize}
\end{lemma}

\begin{proof}
	To prove the first item, note that since $0^i(\sigma+1)0^jw$ is maximal among its rotations, the same holds for $0^{i+1+j}w$. Thus, we need to show that $0^{i+1+j}w\in C_k$.

	Take a maximal sequence of cycles that begins in $C_m$ and let $C_{m+r}$ be the last cycle in this sequence. By lemma \ref{sequence}, $key(C_{m+r})=0^i(\sigma+1+r)0^jw$. By the parentheses property, $C_{m+r}$ is immediately embedded in $C_k$. And finally, by the maximality of the sequence of cycles, $nxt(last(C_{m+r}))\in C_k$.

	As a result, we have:
	$$nxt(last(C_{m+r}))=nxt((\sigma+1+r)0^jw0^i)\in\{\ 0^jw0^i0, 0^jw0^i(\sigma+2+r)\}.$$ If $nxt((\sigma+1+r)0^jw0^i)= 0^jw0^i(\sigma+2+r)$, then $last((\sigma+2+r)0^jw0^i)$ which implies $first(0^jw0^i(\sigma+2+r))$. But $first(last(C_{m+r}))$ contradicts the maximality of our sequence of cycles. Hence, $nxt((\sigma+1+r)0^jw0^i)= 0^jw0^i0\in C_k$. Now, $0^{i+1+j}w$ is a rotation of $0^jw0^i0$ thus $0^{i+1+j}w\in C_k$ as required.

	For proving the second item, we note that $last(C_k)=w0^{i+1+j}$. As we have seen, the first element in $C_k$ that follows $C_m$ (and follows $C_{m+r}$) is $nxt(last(C_{m+r}))= 0^jw0^i0$. As a result, if $u\in C_k$ and $last(C_m)<u$, then 
	$$0^jw0^i0\leq u \leq w0^{i+1+j}$$
	which implies that $u=0^{j_2}w0^{i+1+j_1}$
\end{proof}

\begin{corrolary}
	\label{embedding-deletes-zeroes}
	Assume that $C_m$ is $r$-embedded in $C_k$. Write $key(C_m)=uv$ where $u$ is the minimal prefix of $key(C_m)$ that includes $r$ non-zero numbers. Then, $key(C_k)=0^{|u|}v$.
\end{corrolary}

\begin{proof}
	By $r$ invocations of the first item of the previous lemma.
\end{proof}

\begin{lemma}
	\label{first-ordering}
	For every $m\in\N$, $first(C_m)<first (C_{m+1})$.
\end{lemma}

\begin{proof}
	TBD
\end{proof}

\begin{theorem}
	For any word $\tau w$, $\tau w<(\tau+1)w$.
\end{theorem}

\begin{proof}
	Fix an arbitrary cycle $C_m$. We show that if $(\tau+1)w\in C_m$. Then, $\tau w<(\tau+1)w$. We start by showing this fact for $last(C_m)$. Write $key(C_m)=0^l(\sigma+1)w'$ and hence, $last(C_m)=(\sigma+1)w'0^l$ and $first(C_m)=w'0^l(\sigma+1)$. Write $pre=nxt^{-1}$ and note that $pre(w'0^l(\sigma+1))=\sigma w'0^l$. Hence, we get that 
	$$\sigma w' 0^l<(\sigma+1)w'0^l.$$

	Now we deal with the general case in which $(\tau+1)w\neq last(C_m)$. We write 
	$key(C_m)=0^l(\sigma+1)w_1(\tau+1)w_2$, where $$(\tau+1)w=(\tau+1)w_20^l(\sigma+1)w_1.$$ Let $C_k$ be the cycle of $\tau w$. Since every rotation of $\tau w$ is lexicographically smaller than some rotation of $(\tau+1)w$, we have $key(C_k)<_{lex}key(C_m)$. Hence, $k<m$ and by Lemma \ref{first-ordering}, we get 
	$$first(C_k)<first(C_m).$$

	Now, if $last(C_k)<first(C_m)$ then every element of $C_k$ precedes every element of $C_m$ and we are done. Otherwise, by the parentheses property, $C_m$ is embedded in $C_k$. Note that $|(\tau+1)w|_0-|\tau w|_0\in\{0,1\}$. Use corollary \ref{embedding-deletes-zeroes} to conclude that $|(\tau+1)w|_0-|\tau w|_0=1$ and that $C_m$ is immediately embedded in $C_k$. Moreover, note that as $|(\tau+1)w|_0-|\tau w|_0=1$, we have $\tau=0$.

	According to our conclusions, we can write: $key(C_m)=0^l(\sigma+1)w_11w_2$. We write $w_1=0^jw_1'0^i$ and $w_2=0^rw_2'$ where $w_1'$ and $w_2'$ do not start or end with zero. We have,
	$$key(C_m)=0^l(\sigma+1)0^jw_1'0^i10^rw_2'.$$
	Assume for a contradiction that $(\tau+1)w<\tau w$. Recall that $\tau w\in C_k$ and that $C_m$ is immediately embedded in $C_k$, and conclude that $last(C_m)<\tau w\leq last(C_k)$. Therefore, by Lemma \ref{embedding-properties}, $\tau w=0^{j_2}w_1'0^i10^rw_2'0^{l+1+j_1}.$ Thus, we have:
	\begin{equation}
	\label{1}
	0^{j_2}w_1'0^i10^rw_2'0^{l+1+j_1}=0^{r+1}w_2'0^l(\sigma+1)0^jw_1'0^i.
	\end{equation}
	Since also $|0^{j_2}w_1'0^i10^rw_2'0^{l+1+j_1+1}|_1 = |0^{r+1}w_2'0^l(\sigma+1)0^jw_1'0^i|_1$ we get $\sigma+1=1$. Hence,

	\begin{equation}
	\label{2}
	key(C_m)=0^l10^jw_1'0^i10^rw_2'
	\end{equation}

	and Equation \ref{1} can be rewritten as follows:

	\begin{equation}
	\label{3}
	0^{j_2}w_1'0^i10^rw_2'0^{l+1+j_1}=0^{r+1}w_2'0^l10^jw_1'0^i.
	\end{equation}

	For the rest of the proof we assume that $w_1'\neq\varepsilon$ and $w_2'\neq \varepsilon$. The other cases are dealt similarly\footnote{really! I checked !}.
	By deleting the initial and final segments of zeroes, we get from Equation \ref{3},
	\begin{equation}
	\label{4}
	j_2=r+1, \ \ w_1'0^i10^rw_2'=w_2'0^l10^jw_1'.
	\end{equation}
	Now, by Equation \ref{2}, 
	\begin{equation}
	\label{5}
	0^l10^jw_1'0^i10^rw_2'\geq_{lex} 0^i10^rw_2'0^l10^{j}w_1'.
	\end{equation}
	By combining equations \ref{4} and \ref{5}, we get 
	\begin{equation}
	0^l10^j\geq_{lex}0^i10^r.
	\end{equation}
	Hence, $j\leq r$ and in particular, $j_2\leq r$ in contradiction to Equation \ref{4}.
\end{proof}
